---
title: "Danish Kings"
author: "Luna, Nina, STephanie and Andrea"
date: "10/5- 2026"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```


The task here is to load your Danish Monarchs csv into R using the `tidyverse` toolkit, calculate and explore the kings' duration of reign with pipes `%>%` in `dplyr` and plot it over time.

## Load the kings

```{r eval=FALSE}
library(tidyverse)

dir.create("data")
dir.create("figures")

```

```{r eval = TRUE}
#CODE

kings <- read_csv2("data/danishmonarchscsv.csv", na = c("na"))

class(kings)
glimpse(kings)

```


## Calculate the duration of reign for all the kings in your table

Create a new column called  `duration` in the kings dataset, utilizing the `mutate()` function from tidyverse. Check with your group to brainstorm the options.

```{r}
# CODE
view(kings) 

kings <- kings %>%
  mutate(duration = if_else(
    is.na(beg_reg_y) | is.na(end_reg_y),
    NA_real_,                     # If end or start NA then duration NA
    end_reg_y - beg_reg_y       # if not calculate duration
  ))


```

## Calculate the average duration of reign for all rulers

```{r}
# CODE
kings %>%
  summarise(mean(duration, na.rm=TRUE)) # calculates average without missing

avg_duration <- kings %>%
  summarise(avg = mean(duration, na.rm = TRUE)) %>%
  pull(avg) # now the average is saved as a value.
```

## How many and which kings enjoyed a longer-than-average duration of reign?

```{r}
# CODE

kings %>%
  filter(duration >avg_duration) %>% #filter only above average
  count() # counts how many rules are above average duration

```


## How many days did the three longest-ruling monarchs rule? 

```{r}
# CODE

top_three <- kings %>%
  arrange(desc(duration)) %>%     # arranges from high to low duration
  slice_head(n = 3)               # picks the 3 highest
  top_three_days_simple <- top_three %>%
  mutate(Days = duration * 365L)
  
view(top_three)
  
#With leapyear taken into consideration: 
  count_leaps <- function(start_year, end_year) { # creates function
  yrs <- seq(start_year, end_year - 1)              # creates sequence counting whole years in duration
  is_leap <- (yrs %% 400 == 0) | (yrs %% 4 == 0 & yrs %% 100 != 0) # test if year is leapyear
  sum(is_leap, na.rm = TRUE) # counts leapyears 
}
# calculating for the three monarcs with leap years

  top_three_days_bonus <- top_three %>%
  rowwise() %>% # does it row by row
  mutate(
    LeapDays = count_leaps(beg_reg_y, end_reg_y), #counts leap years
    Days_adjusted = duration * 365L + LeapDays      # years in days + leap days
  ) %>%
  ungroup() # end of rowwise

select(top_three_days_bonus, name, Days_adjusted) # shows the results

```


## Challenge: Plot the kings' duration of reign through time


```{r}
#CODE

# creates the midyear column
kings <- kings %>%
  mutate(midyear = beg_reg_y + (duration / 2)      # calculates reign half way point
  )

# removes missing/ na observations
kings_clean <- kings %>%  
  filter(!is.na(midyear), !is.na(duration)) # only keep columns where both midyear and duration is not NA

# creates plot
library(ggplot2)
Duration_of_Reign_Over_Time <- ggplot(kings_clean, aes(x = midyear, y = duration)) +
  geom_point(size = 3, alpha = 0.7) + # creates points in plot
  geom_smooth(method = "loess", se = FALSE, color = "blue") + # creates the line in plot
  theme_bw(base_size = 16) +   # sort/hvid og større font
  labs(
    title = "Duration of Reign Over Time",
    x = "Midyear of Reign",
    y = "Duration of Reign (years)"
  )
 ggsave("figures/Duration_of_Reign_Over_Time.png", plot = Duration_of_Reign_Over_Time)

```